---
title: "HW5"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load required packages:
```{r load}
library(rvest)
library(factoextra)
library(fastICA)
library(graphics)
library(ggplot2)
library(scatterplot3d)
library(nFactors)
```

# Q1 Parkinson's Disease example
Import and prepare parkinson disease data:
```{r import}
wiki_url = read_html("http://wiki.socr.umich.edu/index.php/SOCR_Data_PD_BiomedBigMetadata")
html_nodes(wiki_url, "#content")
pd_data = html_table(html_nodes(wiki_url, "table")[[1]])
pd_data = data.frame(pd_data)
```
Change the "Dx" to numerical, and delete the time and case no. columns:
```{r prepare}
pd_data$Dx = gsub("PD", 1, pd_data$Dx)
pd_data$Dx = gsub("HC", 0, pd_data$Dx)
pd_data$Dx = gsub("SWEDD", 0, pd_data$Dx)
pd_data$Dx = as.numeric(pd_data$Dx)

pd_0 = pd_data[, -c(1, 33)]
summary(pd_0)
```

PCA
```{r PCA}
pd.pca = princomp(pd_0, cor = T)
summary(pd.pca)
```
From the cumulative proportion of variance, 26 principal components can explain over 90% of the variance, so PCA does not work well.

Plot the principal components and the biplot:
```{r pca plot}
plot(pd.pca)
biplot(pd.pca)
```

SVD
```{r SVD}
df = nrow(pd_0) - 1
zvars = scale(pd_0)
z.svd = svd(zvars)
```
d from SVD and eigen values from PCA are the same:
```{r d and eigenval}
z.svd$d/sqrt(df)
pd.pca$sdev
```
v from SVD and eigen vectors from PCA are the same:
```{r v and eigenvec}
z.svd$v[1:10,1:10]
pd.pca$loadings[1:10,1:10]
```

ICA
First we calculate the correlation of the data matrix:
```{r cor}
cor(pd_0)[1:10,1:10]
``` 
The data show low correlation, so not suitable for ICA. But as a practice:
```{r ICA}
pd.ica = fastICA(pd_0, 2, alg.typ = "parallel", fun = "logcosh", alpha = 1, 
                 method = "C", row.norm = T, maxit = 200, 
                 tol = 0.0001)
```
We can then plot the two ICA components and show the new correlation matrix of the two components:
```{r ICA plot}
plot(pd.ica$S, main = "ICA components")
cor(pd.ica$S)
```

FA
```{r FA}
pd.fa = factanal(pd_0, 19, rotation = "varimax")
pd.fa
```
We can see that 19 factors pass the chi square test for sufficient number of factors at 0.78 level.

# Q2 Allometric Relations in Plants example

Load data, using the "Boreal/temperate Larix forest" dataset:
```{r}
wiki_url = read_html("http://wiki.stat.ucla.edu/socr/index.php/SOCR_Data_Dinov_032708_AllometricPlanRels")
html_nodes(wiki_url, "#content")
pl_data = html_table(html_nodes(wiki_url, "table")[[2]])
pl_data = data.frame(pl_data)
```
Delete the sites column since the information is included in the altitudes and latitudes. Also delete the "Born" columns since all have the same content.
```{r}
pl_raw = pl_data
pl_data = pl_data[,-c(1,5)]
```
Generate a data summary:
```{r}
head(pl_data)
summary(pl_data)
```
Apply prcomp and report the rotations (scores). Data is scaled.
```{r}
pl.pca = prcomp(pl_data, scale. = T)
summary(pl.pca)
pl.pca
```
Display screen plot:
```{r}
plot(pl.pca)
```
Noticed that 4 principal components explain 98.8% of the variance, so choose 4 pc. Now employ a bootstrap test:
```{r}
set.seed(2017)
num_boot = 1000
bootstrap_it = function(i) {
  data_resample = pl_data[sample(1:nrow(pl_data), nrow(pl_data), replace=TRUE),] 
  p_resample = prcomp(data_resample, scale. = T) 
  return(sum(p_resample$sdev[1:4]^2)/sum(p_resample$sdev^2))
  }
pco = data.frame(per=sapply(1:num_boot, bootstrap_it)) 
quantile(pco$per, probs = c(0.025,0.975))  # specify 95-th % Confidence Interval

corpp = sum(pl.pca$sdev[1:4]^2)/sum(pl.pca$sdev^2)
plot = ggplot(pco, aes(x=pco$per)) +
  geom_histogram(bins = 50) + geom_vline(xintercept=corpp, color='yellow')+ 
  labs(title = "Percent Var Explained by the first 4 PCs") +
  theme(plot.title = element_text(hjust = 0.5))+
  labs(x='perc of var')
show(plot)
```
Apply factoextra to draw biplot and grouped by Province/Sites:
```{r}
qualit_vars = as.factor(pl_raw$Province.Sites)
fviz_pca_biplot(pl.pca, axes = c(1, 2), geom = c("point", "text"),
  col.ind = "black", col.var = "steelblue", label = "all",
  invisible = "none", repel = T, habillage = qualit_vars, 
  palette = NULL, addEllipses = TRUE, title = "PCA - Biplot")
```
Perform SVD and compare the results of PCA.
SVD:
```{r}
df = nrow(pl_data) - 1
zvars = scale(pl_data)
z.svd = svd(zvars)
```
d from SVD and eigen values from PCA are the same:
```{r}
z.svd$d/sqrt(df)
```
```{r}
pl.pca$sdev
```
v from SVD and eigen vectors from PCA are the same:
```{r}
z.svd$v
```
```{r}
pl.pca$rotation
```
Use these three variables L, M, D to perform ICA and show pair plots before ICA and after ICA. 
Scatter plot before ICA:
```{r}
pl.sub = pl_data[,4:6]
scatterplot3d(pl.sub)
```
Scatter plot after ICA:
```{r}
pl.ica = fastICA(pl.sub, 3, alg.typ = "parallel", fun = "logcosh", alpha = 1, 
                 method = "C", row.norm = T, maxit = 200, tol = 0.0001)
scatterplot3d(pl.ica$S, main = "ICA components")
```
We can see that ICA components are more indepedent except one pair.
The correlation of PCs from PCA are:
```{r}
plsub.pca = prcomp(pl.sub, scale. = T)
cor(plsub.pca$rotation)
```
While the correlation of the components from ICA are:
```{r}
cor(pl.ica$S)
```
PCA and ICA both have 2 components that are very correlated, implying that only 2 components are enough.

Factor analysis
Use require(nFactors) to determine the number of the factors and show a scree plot as stated in notes
```{r}
ev = eigen(cor(pl_data)) # get eigenvalues
ap = parallel(subject=nrow(pl_data), var=ncol(pl_data), rep=100, cent=.05)
nS = nScree(x=ev$values, aparallel=ap$eigen$qevpea)
summary(nS)
```
```{r}
plotnScree(nS)
```
So the number of factors is determined to be 2.
Use factanal() to apply FA. When rotation =  varimax:
```{r}
pl.favar = factanal(pl_data, 2, rotation = "varimax")
pl.favar
```
When rotation = promax:
```{r}
pl.fapro = factanal(pl_data, 2, rotation = "promax")
pl.fapro
```
Loadings are shown in both fa results. We get similar results.
Visualizing the results (use result from varimax):
```{r}
load = pl.favar$loadings
plot(load, type="n") 
text(load, labels=colnames(pl_data), cex=.7)
```
We can see that using PCA, ICA or FA, we can reduce the data to 2 components.
