---
title: "HW 6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load required packages:
```{r}
library(rvest)
library(factoextra)
library(graphics)
library(ggplot2)
library(e1071)
library(class)
library(reshape2)
library(GGally)
library(gmodels)
```

# Q1 Traumatic Brain injury
Use the kNN algorithm to provide a classification of the TBI SOCR data. DIchotomize the field.gcs outcome variable by field.gcs>=7. Determine an appropriate kk, train and evaluate the performance of the classification model on the data. Report some model quality statistics for a couple of different values of kk, and use these to rank-order (and perhaps plot the classification results of) the models.

First load the TBI SOCR data:
```{r}
wiki_url = read_html("http://wiki.socr.umich.edu/index.php/SOCR_Data_N46_TBI_ROI_Volumes")
html_nodes(wiki_url, "#content")
tbi = html_table(html_nodes(wiki_url, "table")[[1]])
tbi = data.frame(tbi)
colnames(tbi) = tbi[1,]
tbi = tbi[-1,]
```

```{r}
str(tbi)
```

Need to convert them into numeric:
```{r}
tbi = as.data.frame(apply(tbi[,1:18],2, function(x) as.numeric(x)))
```

Then dichotomize the field.gcs outcome variable by field.gcs>=7. Small than 7 will be 0 and greater than 7 will be 1:
```{r}
tbi$field.gcs[which(tbi$field.gcs < 7)] = 0
tbi$field.gcs[which(tbi$field.gcs >= 7)] = 1
tbi$field.gcs = as.factor(tbi$field.gcs)
```

Delete the ID column, split the training and testing sets, and define x and y:
```{r}
tbi = tbi[,-1]
tbi_train = tbi[1:23,]
tbi_test = tbi[24:31,]
tbi_train.x = tbi_train[,-3]
tbi_test.x = tbi_test[,-3]
tbi_train.y = tbi_train[,3]
tbi_test.y = tbi_test[,3]
```

Standardize data:
```{r}
tbi_test.xz = scale(tbi_test.x)
tbi_train.xz = scale(tbi_train.x)
```

Determine an appropriate kk:
```{r}
knntuning = tune.knn(x = tbi_train.xz, y = tbi_train.y, k = 1:6)
summary(knntuning)
```

Choose k = 1.
Train and evaluate the performance of the classification model on the data:
```{r}
tbi.knn1 = knn(train=tbi_train.xz, test=tbi_test.xz, cl=tbi_train.y, k=1)
CrossTable(x=tbi_test.y, y=tbi.knn1, prop.chisq = F)
```

Report some model quality statistics for a couple of different values of k.
k = 2:
```{r}
tbi.knn2 = knn(train=tbi_train.xz, test=tbi_test.xz, cl=tbi_train.y, k=2)
CrossTable(x=tbi_test.y, y=tbi.knn2, prop.chisq = F)
```

k = 3:
```{r}
tbi.knn3 = knn(train=tbi_train.xz, test=tbi_test.xz, cl=tbi_train.y, k=3)
CrossTable(x=tbi_test.y, y=tbi.knn3, prop.chisq = F)
```

k = 4:
```{r}
tbi.knn4 = knn(train=tbi_train.xz, test=tbi_test.xz, cl=tbi_train.y, k=4)
CrossTable(x=tbi_test.y, y=tbi.knn4, prop.chisq = F)
```

 = 5:
```{r}
tbi.knn5 = knn(train=tbi_train.xz, test=tbi_test.xz, cl=tbi_train.y, k=5)
CrossTable(x=tbi_test.y, y=tbi.knn5, prop.chisq = F)
```

Visualize the error rate against the value of k (modified from the code in the note):
```{r}
cv_partition <- function(y, num_folds = 10, seed = NULL) {
  if(!is.null(seed)) {
    set.seed(seed)
  }
  n <- length(y)

  folds <- split(sample(seq_len(n), n), gl(n = num_folds, k = 1, length = n))
  folds <- lapply(folds, function(fold) {
    list(
      training = which(!seq_along(y) %in% fold),
      test = fold
    )
  })
  names(folds) <- paste0("Fold", names(folds))
  return(folds)
}

# Generate 5-folds of the data
folds = cv_partition(tbi_train.y, num_folds = 5)

# Define a trainingset_CV_error calculation function
train_cv_error_tbi = function(K) {
  #Train error
  knnbt = knn(train = tbi_train.xz, test = tbi_train.xz, 
                 cl = tbi_train.y, k = K)
  train_error = mean(knnbt != tbi_train.y)

  #CV error
  cverrbt = sapply(folds, function(fold) {
    mean(tbi_train.y[fold$test] != knn(train = tbi_train.xz[fold$training,], cl = tbi_train.y[fold$training], 
                                       test = tbi_train.xz[fold$test,], k=K))
    }
  )

  cv_error = mean(cverrbt)

  #Test error
  knn.test = knn(train = tbi_train.xz, test = tbi_test.xz, 
            cl = tbi_train.y, k = K)
  test_error = mean(knn.test != tbi_test.y)
  return(c(train_error, cv_error, test_error))
}

k_err = sapply(1:15, function(k) train_cv_error_tbi(k))
df_errs = data.frame(t(k_err), 1:15)
colnames(df_errs) = c('Train', 'CV', 'Test', 'K')

dataL <- melt(df_errs, id="K")
ggplot(dataL, aes_string(x="K", y="value", colour="variable",
   group="variable", linetype="variable", shape="variable")) +
   geom_line(size=0.8) + labs(x = "Number of nearest neighbors (k)",
           y = "Classification error",
           colour="", group="",
           linetype="", shape="") +
  geom_point(size=2.8)
```

# Q2 Parkinson’s Disease
Import data:
```{r}
setwd('C:/Users/Xylthx/Desktop/HS650/HW6')
pd = read.csv('05_PPMI_top_UPDRS_Integrated_LongFormat.csv')
```
# 2.1 KNN Classification in a High Dimension Space
Process data: delete the Index and FID_IID VisitID column; convert the response variable ResearchGroup to bipolar factor(consider SWEDD as disease).
```{r}
pd = subset(pd, select = -c(Index,FID_IID,VisitID))
pd$ResearchGroup[which(pd$ResearchGroup == 'SWEDD')] = 'PD'
pd$ResearchGroup = factor(pd$ResearchGroup)
table(pd$ResearchGroup)
```

Detect NA values:
```{r}
anyNA(pd)
```

So no NA in the data frame.
Summarize the dataset.
Use str:
```{r}
str(pd)
```

Use summary:
```{r}
summary(pd)
```

Use cor:
```{r}
cor(pd[,-31])[1:10,1:10]
```

Use ggpairs:
```{r}
ggpairs(pd[,1:5])
```

Data Transformation: scale/normalize the data: log(x-min(x)) and discretize either 0 or 1 (smaller than mean is 0 and others is 1).
```{r}
logscale = function(x){
  log(x-min(x))
}

pd.xs = pd[,-31]

for(i in c(1:67)){
  a = logscale(pd.xs[,i])
  pd.xs[which(a < mean(a[a!=-Inf])),i] = 0
  pd.xs[which(a >= mean(a[a!=-Inf])),i] = 1
}

summary(pd.xs)
```

Randomly partition the data into training and testing sets: use set.seed and random sample, train:test=2:1.
```{r}
set.seed(1)
trind = sample(1:661, 441, replace = F)
pd_train.xs = pd.xs[trind,]
pd_test.xs = pd.xs[-trind,]
pd_train.y = pd[trind,31]
pd_test.y = pd[-trind,31]
```

Select an optimized k for each of the scaled data above: Show an error plot for k including three lines: train error, cross validation error and test error.
```{r}
# Generate 10-folds of the data
folds = cv_partition(pd_train.y, num_folds = 10)

# Define a trainingset_CV_error calculation function
train_cv_error_pd = function(K) {
  #Train error
  knnpd = knn(train = pd_train.xs, test = pd_train.xs, 
                 cl = pd_train.y, k = K)
  train_error = mean(knnpd != pd_train.y)

  #CV error
  cverrpd = sapply(folds, function(fold) {
    mean(pd_train.y[fold$test] != knn(train = pd_train.xs[fold$training,], cl = pd_train.y[fold$training], 
                                       test = pd_train.xs[fold$test,], k=K))
    }
  )

  cv_error = mean(cverrpd)

  #Test error
  knn.test = knn(train = pd_train.xs, test = pd_test.xs, 
            cl = pd_train.y, k = K)
  test_error = mean(knn.test != pd_test.y)
  return(c(train_error, cv_error, test_error))
}

k_err = sapply(1:30, function(k) train_cv_error_pd(k))
df_errs = data.frame(t(k_err), 1:30)
colnames(df_errs) = c('Train', 'CV', 'Test', 'K')

dataL <- melt(df_errs, id="K")
ggplot(dataL, aes_string(x="K", y="value", colour="variable",
   group="variable", linetype="variable", shape="variable")) +
   geom_line(size=0.8) + labs(x = "Number of nearest neighbors (k)",
           y = "Classification error",
           colour="", group="",
           linetype="", shape="") +
  geom_point(size=2.8)
```

Select k = 3.
Hypothesize that for training set, generally as k increase the error increases, but there is a optimal k in the middle. Best k for the trainig set does not really mean the best k for the model, as when k is small the model tends to overfit. We can check this with tune.knn:
```{r}
knntuning = tune.knn(x = pd_train.xs, y = pd_train.y, k = 1:30)
summary(knntuning)
```

The errors of the training and testing sets are shown in the figure above. We can see that when k is small it will overfit and when k is large the performance of the model is bad.

Then try above again but for a smaller subset of the data frame:
```{r}
pdsub = pd[,c(31,40,46,47,53,54,60)]

pdsub.xs = pdsub[,-1]

for(i in c(1:6)){
  a = logscale(pdsub.xs[,i])
  pdsub.xs[which(a < mean(a[a!=-Inf])),i] = 0
  pdsub.xs[which(a >= mean(a[a!=-Inf])),i] = 1
}

summary(pdsub.xs)
```

```{r}
set.seed(1)
trind = sample(1:661, 441, replace = F)
pdsub_train.xs = pdsub.xs[trind,]
pdsub_test.xs = pdsub.xs[-trind,]
pdsub_train.y = pdsub[trind,1]
pdsub_test.y = pdsub[-trind,1]

# Generate 10-folds of the data
folds = cv_partition(pdsub_train.y, num_folds = 10)

train_cv_error_sub = function(K) {
  #Train error
  knnpd = knn(train = pdsub_train.xs, test = pdsub_train.xs, 
                 cl = pdsub_train.y, k = K)
  train_error = mean(knnpd != pdsub_train.y)

  #CV error
  cverrpd = sapply(folds, function(fold) {
    mean(pdsub_train.y[fold$test] != knn(train = pdsub_train.xs[fold$training,], 
                                         cl = pdsub_train.y[fold$training], 
                                       test = pdsub_train.xs[fold$test,], k=K))
    }
  )

  cv_error = mean(cverrpd)

  #Test error
  knn.test = knn(train = pdsub_train.xs, test = pdsub_test.xs, 
            cl = pdsub_train.y, k = K)
  test_error = mean(knn.test != pdsub_test.y)
  return(c(train_error, cv_error, test_error))
}

k_err = sapply(1:30, function(k) train_cv_error_sub(k))
df_errs = data.frame(t(k_err), 1:30)
colnames(df_errs) = c('Train', 'CV', 'Test', 'K')

dataL <- melt(df_errs, id="K")
ggplot(dataL, aes_string(x="K", y="value", colour="variable",
   group="variable", linetype="variable", shape="variable")) +
   geom_line(size=0.8) + labs(x = "Number of nearest neighbors (k)",
           y = "Classification error",
           colour="", group="",
           linetype="", shape="") +
  geom_point(size=2.8)
```

```{r}
knntuning = tune.knn(x = pd_train.xs, y = pd_train.y, k = 1:30)
summary(knntuning)
```

Compared to the full data, a slightly smaller k is more optimal.
The analysis above is only done for a log-transformed data after a discretization. Other transformations can be done and the code is provided below:
```{r}
# pd_norm.x = scale(pd[,-31]) # normalization
# pd_log.x = logscale(pd[,-31]) # log-transformation
# A binary transformation can also be done with a for loop with a ifelse inside.
```

















